Dominance Regions for Rank Two Cluster Algebras

### 2022-04-25 dominance_regions.tex
Better still
Edits

  When $bc\le3$, the cluster algebra $\cA(b,c)$ will be of finite-type and cluster monomials form its only pointed basis.
  We will therefore assume that $bc\ge4$.
  In this case we give an explicit description of the dominance relation among $\bfg$-vectors.
  Specifically, we show that the $\bfg$-vector $\lambda$ dominates the collection of $\bfg$-vectors of the form $\lambda+(b \alpha_0 ,c \alpha_1)$, $\alpha_0,\alpha_1\in\ZZ$, inside its \emph{dominance polygon} $\cP_\lambda$.
  %In this case, we compute the possible transitions between any two pointed bases.
  %The collection of $\bfg$-vectors which can appear in the deformation of a pointed basis is the \emph{dominance region}.
  \begin{theorem}
    \label{th:dominance inequalities}
    The dominance polygon $\cP_\lambda$ pointed at $\lambda=(\lambda_0,\lambda_1)$ is the region consisting of those $\mu\in\RR^2$ satisfying $\mu_0 \leq \lambda_0, \mu_1 \geq\lambda_1$, and the following inequalities:
    {
      \everymath={\displaystyle}
      \def\arraystretch{2.8}
      \[
        \begin{array}{rcccl}
          0 & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)+(\mu_1-\lambda_1) & \leq & \max\left(-c\lambda_0-\frac{b c+\sqrt{b c (b c-4)}}{2b}b\lambda_1,0\right)
          \\
          \min\left(-c\lambda_0,0\right) & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)-(\mu_1-\lambda_1) & \leq & 0
          \\
          0 & \leq &  -(\mu_0-\lambda_0)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(\mu_1-\lambda_1) & \leq & \max\left(\frac{b c+\sqrt{b c (b c-4)}}{2c}c\lambda_0+b\lambda_1,0\right)
          \\
          \min\left(b \lambda_1,0\right) & \leq & (\mu_0-\lambda_0) - \frac{b c-\sqrt{b c (b c-4)}}{2 c} (\mu_1-\lambda_1) & \leq & 0
        \end{array}
      \]
    }
  \end{theorem}

@@

  Write $\cI \subset \RR^2$ for the \emph{imaginary cone} (positively) spanned by the vectors $\big(2b,-bc\pm\sqrt{bc(bc-4)}\big)$.
  \begin{corollary}
    \label{cor:dominance vertices}
    There are six classes of dominance polytopes.
    \begin{enumerate}
      \item If $\lambda$ lies outside of $\cI$, i.e. if it is the $\bfg$-vector of a cluster monomial, then $\cP_\lambda$ is the point $\lambda$. 
      \item If $\lambda$ lies in the interior of the cone spanned by the vectors $(2b,-bc-\sqrt{bc(bc-4)})$ and $(2,-c)$, then $\cP_\lambda$ is the trapezoid with corner vertices $\lambda$, $-\frac{bc+\sqrt{bc(bc-4)}}{2c}\big(\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1,0\big)$, $\big(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1\big)$, and $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}\big(-(bc-2)\lambda_0-b\lambda_1,c\lambda_0+2\lambda_1\big)$.
      \item If $\lambda$ lies on the ray spanned by $(2,-c)$, say $\lambda=(2\ell,-c\ell)$, then $\cP_\lambda$ is the triangle with corner vertices $\lambda$, $\big(2\ell-\frac{bc+\sqrt{bc(bc-4)}}{2}\ell,0\big)$, and $\big(0,\frac{\sqrt{bc(bc-4)}}{b}\ell\big)$.
      \item If $\lambda$ lies in the interior of the cone spanned by the vectors $(2,-c)$ and $(b,-2)$, then $\cP_\lambda$ is the kite with corner vertices $\lambda$, $\big(\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1,0\big)$, $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}(2\lambda_0+b\lambda_1,c\lambda_0+2\lambda_1)$, and $\big(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1\big)$.
      \item If $\lambda$ lies on the ray spanned by $(b,-2)$, say $\lambda=(b\ell,-2\ell)$, then $\cP_\lambda$ is the triangle with corner vertices $\lambda$, $\big(-\frac{\sqrt{bc(bc-4)}}{c}\ell,0\big)$, and $\big(0,\frac{bc+\sqrt{bc(bc-4)}}{2}\ell-2\ell\big)$.
      \item If $\lambda$ lies in the interior of the cone spanned by the vectors $(b,-2)$ and $(2b,-bc+\sqrt{bc(bc-4)})$, then $\cP_\lambda$ is the trapezoid with corner vertices $\lambda$, $\big(\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1,0\big)$, $-\frac{bc+\sqrt{bc(bc-4)}}{2b}\big(0,\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1\big)$, and $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}\big(2\lambda_0+b\lambda_1,-c\lambda_0-(bc-2)\lambda_1\big)$.
    \end{enumerate}
  \end{corollary}

@@

  Given a Laurent polynomial in $\ZZ[x_0^{\pm1},x_1^{\pm1}]$, its \emph{support} is the set of its exponent vectors inside $\ZZ^2$.
  \begin{theorem}
    \label{th:maximum support}
    For a $\bfg$-vector $\lambda=(\lambda_0,\lambda_1)\in\ZZ^2$ the support for a pointed basis vector $x_\lambda$ is contained in the (possibly degenerate) lattice quadrilateral $\cS_\lambda$ given as follows:
    \begin{enumerate}
      \item If $\lambda$ lies outside of $\cI$, there are the following possibilities:
        \begin{enumerate}
          \item If $0\le\lambda_0,\lambda_1$, then $\cS_\lambda$ is just the point $\lambda$.
          \item If $\lambda_1 < 0$ and $0\le\lambda_0+b\lambda_1$, then $\cS_\lambda$ is the segment joining $\lambda$ and $(\lambda_0+b\lambda_1,\lambda_1)$.
          \item If $\lambda_0 < 0$ and $0\le\lambda_1$, then $\cS_\lambda$ is the segment joining $\lambda$ and $(\lambda_0,-c\lambda_0+\lambda_1)$.
          \item If $\lambda_0,\lambda_1 < 0$, then $\cS_\lambda$ has corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $(\lambda_0,-c\lambda_0+\lambda_1)$.
          \item If $\lambda_0 > 0$, $\lambda_0+b\lambda_1 < 0$, and $-c\lambda_0-(bc-1)\lambda_1\le 0$, then $\cS_\lambda$ has corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $\big((bc+1)\lambda_0-b^2c\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$.
          \item If $\lambda_0 > 0$, $\lambda_0+b\lambda_1 < 0$, and $0 < -c\lambda_0-(bc-1)\lambda_1$, then $\cS_\lambda$ has corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $(0,0)$.
        \end{enumerate}
      \item If $\lambda$ lies inside of $\cI$, then $\cS_\lambda$ has corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}\big(2\lambda_0+b\lambda_1,-c\lambda_0-(bc-2)\lambda_1\big)$.
    \end{enumerate}
  \end{theorem}

@@

  Every pointed basis element for $\cA(b,c)$ admits an opposite $\bfg$-vector arising by interchanging the roles of $b$, $c$ and $x_0$, $x_1$ in \eqref{eq:pointed}. 
  \begin{lemma}
    Given a $\bfg$-vector $\lambda$, the opposite $\bfg$-vector $\lambda'$ is obtained as follows:
    \begin{itemize}
      \item if $\lambda_0,\lambda_1\ge0$, then $\lambda'=\lambda$;
      \item if $\lambda_1\ge0$ and $\lambda_0<0$, then $\lambda'=(\lambda_0,-c\lambda_0+\lambda_1)$;
      \item if $\lambda_0>0$ and $\lambda_0+b\lambda_1>0$, then $\lambda'=(\lambda_0+b\lambda_1,\lambda_1)$;
      \item otherwise, $\lambda'=(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1)$.
    \end{itemize}
  \end{lemma}

@@

  The support $\cG_\lambda$ of the greedy basis element of $\cA(b,c)$ with $\bfg$-vector $\lambda$ is well-known \cite{LLZ,cgmmrsw}.
  This is indicated by the solid and dashed lines in Theorem~\ref{th:maximum support} where dashed lines indicate points that are excluded from the support (note that the white dot does not indicate that this point is excluded).
  For our purposes, it is enough to observe the following closure property for the greedy support.
  \begin{definition}
    A subset $\cS\subset\RR^2$ is called \emph{min-convex} if $\lambda,\mu\in\cS$ implies the segments joining $\lambda$ and $\mu$ to $\big(\min(\lambda_0,\mu_0),\min(\lambda_1,\mu_1)\big)$ are contained in $\cS$.
  \end{definition}
  In particular, (an upper bound for) the greedy support can be found using the $\bfg$-vector $\lambda$ indicated by a black dot and the opposite $\bfg$-vector $\lambda'$ indicated by a white dot in Theorem~\ref{th:maximum support}.
  Write $\overline{\cG}_\lambda$ for the cone of the min-convex closure of $\{\lambda,\lambda'\}$, that is $\overline{\cG}_\lambda$ contains the segments joining $\lambda$ and $\lambda'$ with $\big(\min(\lambda_0,\lambda'_0),\min(\lambda_1,\lambda'_1)\big)$ and for any $\mu\in\overline{\cG}_\lambda$ we have $t\mu\in\overline{\cG}_\lambda$ for $0\le t\le 1$.
  Clearly, $\cG_\lambda\subset\overline{\cG}_\lambda$.

  As we saw above, for any $\bfg$-vector $\mu\in\cP_\lambda$ its opposite $\bfg$-vector $\mu'$ is also contained in $\cS_\lambda$.
  But $\cS_\lambda$ is min-convex and closed under downward scaling, i.e. for any $\mu\in\cS_\lambda$ we have $t\mu\in\cS_\lambda$ for $0\le t\le 1$.
  It follows that $\overline{\cG}_\mu\subset\cS_\lambda$ for any $\mu\in\cP_\lambda$.

>>>>

  When $bc\le3$, the cluster algebra $\cA(b,c)$ will be of finite-type and cluster monomials form its only pointed basis.
  We will therefore assume that $bc\ge4$.
  In this case we give an explicit description of the dominance relation among $\bfg$-vectors.
  Specifically, we show that the $\bfg$-vector $\lambda$ dominates the collection of $\bfg$-vectors of the form $\lambda+(b \alpha_0 ,c \alpha_1)$, $\alpha_0,\alpha_1\in\ZZ$, inside its \emph{dominance region} $\cP_\lambda$.
  %In this case, we compute the possible transitions between any two pointed bases.
  %The collection of $\bfg$-vectors which can appear in the deformation of a pointed basis is the \emph{dominance region}.
  \begin{theorem}
    \label{th:dominance inequalities}
    The dominance region $\cP_\lambda$ pointed at $\lambda=(\lambda_0,\lambda_1)$ is the region consisting of those $\mu\in\RR^2$ satisfying $\mu_0 \leq \lambda_0, \mu_1 \geq\lambda_1$, and the following inequalities:
    {
      \everymath={\displaystyle}
      \def\arraystretch{2.8}
      \[
        \begin{array}{rcccl}
          0 & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)+(\mu_1-\lambda_1) & \leq & \max\left(-c\lambda_0-\frac{b c+\sqrt{b c (b c-4)}}{2b}b\lambda_1,0\right)
          \\
          \min\left(-c\lambda_0,0\right) & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)-(\mu_1-\lambda_1) & \leq & 0
          \\
          0 & \leq &  -(\mu_0-\lambda_0)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(\mu_1-\lambda_1) & \leq & \max\left(\frac{b c+\sqrt{b c (b c-4)}}{2c}c\lambda_0+b\lambda_1,0\right)
          \\
          \min\left(b \lambda_1,0\right) & \leq & (\mu_0-\lambda_0) - \frac{b c-\sqrt{b c (b c-4)}}{2 c} (\mu_1-\lambda_1) & \leq & 0
        \end{array}
      \]
    }
  \end{theorem}

@@

  Write $\cI \subset \RR^2$ for the \emph{imaginary cone} (positively) spanned by the vectors $\big(2b,-bc\pm\sqrt{bc(bc-4)}\big)$.
  \begin{corollary}
    \label{cor:dominance vertices}
    There are six classes of dominance regions, these are (possibly degenerate) polygons given as follows.
    \begin{enumerate}
      \item If $\lambda$ lies outside of $\cI$, i.e. if it is the $\bfg$-vector of a cluster monomial, then $\cP_\lambda$ is the point $\lambda$. 
      \item If $\lambda$ lies in the interior of the cone spanned by the vectors $(2b,-bc-\sqrt{bc(bc-4)})$ and $(2,-c)$, then $\cP_\lambda$ is the trapezoid with corner vertices $\lambda$, $-\frac{bc+\sqrt{bc(bc-4)}}{2c}\big(\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1,0\big)$, $\big(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1\big)$, and $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}\big(-(bc-2)\lambda_0-b\lambda_1,c\lambda_0+2\lambda_1\big)$.
      \item If $\lambda$ lies on the ray spanned by $(2,-c)$, say $\lambda=(2\ell,-c\ell)$, then $\cP_\lambda$ is the triangle with corner vertices $\lambda$, $\big(2\ell-\frac{bc+\sqrt{bc(bc-4)}}{2}\ell,0\big)$, and $\big(0,\frac{\sqrt{bc(bc-4)}}{b}\ell\big)$.
      \item If $\lambda$ lies in the interior of the cone spanned by the vectors $(2,-c)$ and $(b,-2)$, then $\cP_\lambda$ is the kite with corner vertices $\lambda$, $\big(\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1,0\big)$, $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}(2\lambda_0+b\lambda_1,c\lambda_0+2\lambda_1)$, and $\big(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1\big)$.
      \item If $\lambda$ lies on the ray spanned by $(b,-2)$, say $\lambda=(b\ell,-2\ell)$, then $\cP_\lambda$ is the triangle with corner vertices $\lambda$, $\big(-\frac{\sqrt{bc(bc-4)}}{c}\ell,0\big)$, and $\big(0,\frac{bc+\sqrt{bc(bc-4)}}{2}\ell-2\ell\big)$.
      \item If $\lambda$ lies in the interior of the cone spanned by the vectors $(b,-2)$ and $(2b,-bc+\sqrt{bc(bc-4)})$, then $\cP_\lambda$ is the trapezoid with corner vertices $\lambda$, $\big(\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1,0\big)$, $-\frac{bc+\sqrt{bc(bc-4)}}{2b}\big(0,\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1\big)$, and $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}\big(2\lambda_0+b\lambda_1,-c\lambda_0-(bc-2)\lambda_1\big)$.
    \end{enumerate}
  \end{corollary}

@@

  Given a Laurent polynomial in $\ZZ[x_0^{\pm1},x_1^{\pm1}]$, its \emph{support} is the set of its exponent vectors inside $\ZZ^2$.
  \begin{theorem}
    \label{th:maximum support}
    For a $\bfg$-vector $\lambda=(\lambda_0,\lambda_1)\in\ZZ^2$ the support for a pointed basis vector $x_\lambda$ is contained in the (possibly degenerate) lattice quadrilateral $\cS_\lambda$ given as follows:
    \begin{enumerate}
      \item If $\lambda$ lies outside of $\cI$, then the support of $x_\lambda$ is precisely the points of the form $\lambda+(b \alpha_0 ,c \alpha_1)$, $\alpha_0,\alpha_1\in\ZZ$, inside $\cS_\lambda$ given as follows:
        \begin{enumerate}
          \item If $0\le\lambda_0,\lambda_1$, then $\cS_\lambda$ is just the point $\lambda$.
          \item If $\lambda_1 < 0$ and $0\le\lambda_0+b\lambda_1$, then $\cS_\lambda$ is the segment joining $\lambda$ and $(\lambda_0+b\lambda_1,\lambda_1)$.
          \item If $\lambda_0 < 0$ and $0\le\lambda_1$, then $\cS_\lambda$ is the segment joining $\lambda$ and $(\lambda_0,-c\lambda_0+\lambda_1)$.
          \item If $\lambda_0,\lambda_1 < 0$, then $\cS_\lambda$ has corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $(\lambda_0,-c\lambda_0+\lambda_1)$.
          \item If $\lambda_0 > 0$, $\lambda_0+b\lambda_1 < 0$, and $-c\lambda_0-(bc-1)\lambda_1\le 0$, then $\cS_\lambda$ has corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $\big((bc+1)\lambda_0-b^2c\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$.
          \item If $\lambda_0 > 0$, $\lambda_0+b\lambda_1 < 0$, and $0 < -c\lambda_0-(bc-1)\lambda_1$, then $\cS_\lambda$ has corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $(0,0)$.
        \end{enumerate}
      \item If $\lambda$ lies inside of $\cI$, then $\cS_\lambda$ has corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}\big(2\lambda_0+b\lambda_1,-c\lambda_0-(bc-2)\lambda_1\big)$.
    \end{enumerate}
  \end{theorem}

@@

  Every pointed basis element for $\cA(b,c)$ admits an opposite $\bfg$-vector arising by interchanging the roles of $b$, $c$ and $x_0$, $x_1$ in \eqref{eq:pointed}. 
  One can easily compute the following correspondence.
  \begin{lemma}
    Given a $\bfg$-vector $\lambda$, the opposite $\bfg$-vector $\lambda'$ is obtained as follows:
    \begin{itemize}
      \item if $\lambda_0,\lambda_1\ge0$, then $\lambda'=\lambda$;
      \item if $\lambda_1\ge0$ and $\lambda_0<0$, then $\lambda'=(\lambda_0,-c\lambda_0+\lambda_1)$;
      \item if $\lambda_0>0$ and $\lambda_0+b\lambda_1>0$, then $\lambda'=(\lambda_0+b\lambda_1,\lambda_1)$;
      \item otherwise, $\lambda'=(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1)$.
    \end{itemize}
  \end{lemma}

@@

  The support $\cG_\lambda$ of the greedy basis element of $\cA(b,c)$ with $\bfg$-vector $\lambda$ is well-known \cite{LLZ,cgmmrsw}.
  This is indicated by the solid and dashed lines in Theorem~\ref{th:maximum support} where dashed lines indicate points that are excluded from the support (note that the white dot does not indicate that this point is excluded).
  For our purposes, it is enough to observe the following closure property for the greedy support.
  \begin{definition}
    A subset $\cS\subset\RR^2$ is called \emph{min-convex} if $\lambda,\mu\in\cS$ implies the segments joining $\lambda$ and $\mu$ to $\big(\min(\lambda_0,\mu_0),\min(\lambda_1,\mu_1)\big)$ are contained in $\cS$.
  \end{definition}
  In particular, (an upper bound for) the greedy support can be found using the $\bfg$-vector $\lambda$ indicated by a black dot and the opposite $\bfg$-vector $\lambda'$ indicated by a white dot in Theorem~\ref{th:maximum support}.
  Write $\overline{\cG}_\lambda$ for the downward scaling of the min-convex closure of $\{\lambda,\lambda'\}$, that is $\overline{\cG}_\lambda$ contains the segments joining $\lambda$ and $\lambda'$ with $\big(\min(\lambda_0,\lambda'_0),\min(\lambda_1,\lambda'_1)\big)$ and for any $\mu\in\overline{\cG}_\lambda$ we have $t\mu\in\overline{\cG}_\lambda$ for $0\le t\le 1$.
  Clearly, $\cG_\lambda\subset\overline{\cG}_\lambda$.

  As we saw above, for any $\bfg$-vector $\mu\in\cP_\lambda$ its opposite $\bfg$-vector $\mu'$ is also contained in $\cS_\lambda$.
  But $\cS_\lambda$ is min-convex and closed under downward scaling, i.e. for any $\mu\in\cS_\lambda$ we have $t\mu\in\cS_\lambda$ for $0\le t\le 1$.
  It follows that $\overline{\cG}_\mu\subset\cS_\lambda$ for any $\mu\in\cP_\lambda$ and so the support of any basis element pointed at $\lambda$ is contained in $\cS_\lambda$.
